A New Method of Modelling Tuneable Lasers with Functional Composition

### 2019-05-22 Thesis/Nonlinear/Nonlinear.tex
Made some edits.

In the first experiment the pulse is allowed to equilibrate for 40 loops of the circuit, we then compare the variance, chirp, and phase shift of the two methods. The results of this are shown in Figure \ref{fig:var}. For the most part, we see exceptional agreement between the analytic solution and numerical solution for the variance, and phase shift. However, for $s > 2.4$ the chirp from the simulations seems to erratically vary from the analytic solution. The reason for this is quite a simple one. \\
\begin{figure}[tbp]
\centering
\input{./Figures/LinearPlot}
\caption{Equilibrium energy, and peak power of the pulse as a function of $s$. The thin black line is to highlight that the energy is \emph{not} linearly related to $s$.}
\label{fig:valenergy}
\end{figure}

@@

\section{Nonlinear Model}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_Shape}
\caption{Envelope}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_FT}
\caption{Fourier transform}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_Chirp}
\caption{Chirp}
\end{subfigure}
\caption{Simulation with $s = 0.15$, and $b = 2.1$ after 15 circuits.}
\label{fig:nlstable}
\end{figure}
With the inclusion of the nonlinearity we generally find a similar solution to the linear case. An example of this is shown in Figure \ref{fig:nlstable}. The envelope of the pulse is unsurprisingly Gaussian-esque, however, it is \emph{not} precisely Gaussian and more closely resembles a generalized Gaussian\footnote{A generalized Gaussian has the form $\exp \left(-t^\alpha \right)$, with $\alpha > 2$.}. The fact that the pulse envelope is not Gaussian is further emphasized while examining the Fourier transform of the pulse. If the pulse were a Gaussian, we would expect the Fourier transform to also be a Gaussian \cite{debnath, gradshteyn}. Instead, the magnitude of the Fourier transform has a unique Batman-like shape. This deviation suggests the nonlinearity implants higher frequency oscillations into the pulse---this will be a key observation in the coming subsections. Finally, we shall examine the derivative of the phase---essentially the chirp. Recall in Chapter \ref{chap:linear} the chirp was defined as coefficient of $-\frac{1}{2}iT^2$ in the exponential, by taking the negative of the derivative we would expect a linear function with a slope equal to the chirp. In the nonlinear case, this is what we find for moderate values of $T$. On the other hand, for $|T| > 1$ the relation begins to level off. This is in agreement with experimental results \cite{chen, rothenberg, tomlinson}. \\

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\input{./Figures/StabilityZoom}
\caption{Centralized for more typical $s$, and $b$ values.}
\label{fig:energyzoom}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\input{./Figures/Stability}
\caption{Energy of the pulse within the $s$-$b$ plane.}
\label{fig:energybig}
\end{subfigure}
\caption{Energy of the pulse at equilibrium. The energy is constant along the black lines.}
\label{fig:energy}
\end{figure}

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_Shape}%
\input{./Figures/Unstable_Bad_Shape}
\caption{Envelope}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_FT}%
\input{./Figures/Unstable_Bad_FT}
\caption{Fourier transform}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_Chirp}%
\input{./Figures/Unstable_Bad_Chirp}
\caption{Chirp}
\end{subfigure}
\caption{$s = 0.15$, $b = 2.15$, \textbf{Left:} 11 loops, \textbf{Right:} 16 loops}
\label{fig:break}
\end{figure}

@@

The noise in the energy exhibited for moderate to large values of $b$, and small values of $s$ is a phenomenon called \emph{wave breaking} \cite{agrawal2013, anderson, finot, rothenberg, tomlinson}. Wave breaking is not limited to just optics, wave breaking occurs in areas such as plasmas, transmission lines, and fluid dynamics \cite{rothenberg}. Wave breaking occurs because the pulse begins to interfere with itself in a way called self-phase modulation \cite{agrawal2002, agrawal2013, becker}. Self-phase modulation occurs because the index of refraction is intensity dependent \cite{agrawal2002, becker, rothenberg, silfvast}, which leads to additional chirp across the pulse \cite{agrawal2013, anderson, rothenberg, silfvast}. This in turn causes higher order frequencies to be injected into the pulse \cite{agrawal2013, anderson}. These high frequencies compound with each trip around the cavity becoming parasitic very quickly---Figure \ref{fig:break} highlights this. Notice that the difference between this figure and Figure \ref{fig:nlstable} is a difference in $b$ of $0.05$---this difference could be as small as adding a few centimetres more of fibre between the gain and output coupler. The left figures show the pulse after 11 trips around the cavity, in the Fourier transform it is clear that the contributions from higher frequencies has increased---we obtain similar results as in \cite{anderson, rothenberg}. Additionally, the chirp starts losing its linearity causing it to start becoming unstable; the nature of this instability is in agreement with \cite{anderson, rothenberg}. The parasitic nature of the high frequency contributions is evident by examining the right figures. After 5 additional trips around the cavity, the envelope of the pulse is much more rippled, and the real and imaginary parts become incoherent. Moreover, the Fourier transform has no clear defined shape and has essentially become noise, and the chirp has grown to be highly oscillatory and unstable. Once the pulse has reached a state such as this, there is no turning back---the envelope, Fourier transform, and chirp never reach a steady equilibrium state. \\

\subsection{Convergence}
\begin{figure}[tbp]
\input{./Figures/Step1}
\caption{Error of the pulse (given by \eqref{eq:error}) between iterations $99$ and $100$.}
\label{fig:error}
\end{figure}

\begin{figure}[p]
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step2}
\caption{$\Delta = 2$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step3}
\caption{$\Delta = 3$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step5}
\caption{$\Delta = 5$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step7}
\caption{$\Delta = 7$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step11}
\caption{$\Delta = 11$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step13}
\caption{$\Delta = 13$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step16}
\caption{$\Delta = 16$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step60}
\caption{$\Delta = 60$}
\end{subfigure} \\
\caption{Error for various values of $\Delta$, the colour bar and axis labels have been omitted for clarity, the scales and ranges are the same as in Figure \ref{fig:error}.}
\label{fig:deltaerror}
\end{figure}

@@

The second cause of these errors is because some equilibrium states have a periodicity greater than $1$, we redefine \eqref{eq:error} to be instead
\begin{align}
\textrm{E}_\Delta = \frac{\| |A_i| - |A_{i-\Delta}| \|_2}{\| A_{i-\Delta} \|_2},
\label{eq:deltaerror}
\end{align}
to pick up equilibrium states with period $\Delta$, again with $i = 100$. Figure \ref{fig:deltaerror} shows the error for an assortment of $\Delta$ values. There are several key observations to make, first, for $\Delta = 2$ the kidney-shaped blob in the lower right corner has vanished, thus, this entire region has a periodicity of $2$. Second, for $\Delta = 3$ in the upper right corner a negative triangular-shaped structure emerges. Lastly, $\Delta \in \left\{ 5, 7, 11, 13 \right\}$ may not seem to extract any higher order periods, however, there are in fact very small striations in the lower left. Up until now the $\Delta$ values have been prime since they are the building blocks for composite numbers. However, there are two composite $\Delta$ values of interest---the first being $16$. Since $16 = 2^4$ it will pick up any periodic behaviour with orders of $2, 4, 8, 16$, which are frequently found in bifurcations. Lastly, we chose $\Delta = 60$ for a similar reason: because $60$ is highly divisible---it is an abundant number. This extracts orders such as $12$ or $30$ without having to explicitly run the calculations\footnote{In the case of $\Delta = 60$ a value of $i = 150$ is used to ensure sufficient iterations to converge.}. \\

\begin{figure}[tbp]
\input{./Figures/Min}
\caption{Composite error, \eqref{eq:errorcomp}, of Figure \ref{fig:deltaerror}.}
\label{fig:errorcomp}
\end{figure}

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\input{./Figures/StabilitySwitchZoom}
\caption{Centralized for more typical $s$, and $b$ values.}
\label{fig:switchzoom}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\input{./Figures/StabilitySwitch}
\caption{Energy of the pulse within the $s$-$b$ plane.}
\label{fig:switchbig}
\end{subfigure}
\caption{Energy of the pulse at equilibrium, with the modulation and dispersion blocks switched.}
\label{fig:switch}
\end{figure}

>>>>

In the first experiment the pulse is allowed to equilibrate for 40 loops of the circuit, we then compare the variance, chirp, and phase shift of the two methods. The results of this are shown in Figure \ref{fig:var}. For the most part, we see exceptional agreement between the analytic solution and numerical solution for the variance, and phase shift. However, for $s > 2.4$ the chirp from the simulations seems to erratically vary from the analytic solution. The reason for this is quite a simple one. \\
\begin{figure}[tbp]
\centering
\input{./Figures/LinearPlot}
\caption[Equilibrium energy, and peak power of the pulse as a function of $s$.]{Equilibrium energy, and peak power of the pulse as a function of $s$. The thin black line is to highlight that the energy is \emph{not} linearly related to $s$.}
\label{fig:valenergy}
\end{figure}

@@

\section{Nonlinear Model}
\begin{figure}[tbp]
\centering
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_Shape}
\caption{Envelope}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_FT}
\caption{Fourier transform}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\centering
\input{./Figures/Stable_Chirp}
\caption{Chirp}
\end{subfigure}
\caption[Envelope, Fourier transform, and chirp of the pulse---stable case.]{Simulation with $s = 0.15$, and $b = 2.1$ after 15 circuits.}
\label{fig:nlstable}
\end{figure}
With the inclusion of the nonlinearity we generally find a similar solution to the linear case. An example of this is shown in Figure \ref{fig:nlstable}. The envelope of the pulse is unsurprisingly Gaussian-esque, however, it is \emph{not} precisely Gaussian and more closely resembles a generalized Gaussian\footnote{A generalized Gaussian has the form $\exp \left(-t^\alpha \right)$, with $\alpha > 2$.}. The fact that the pulse envelope is not Gaussian is further emphasized while examining the Fourier transform of the pulse. If the pulse were a Gaussian, we would expect the Fourier transform to also be a Gaussian \cite{debnath, gradshteyn}. Instead, the magnitude of the Fourier transform has a unique Batman-like shape. This deviation suggests the nonlinearity implants higher frequency oscillations into the pulse---this will be a key observation in the coming subsections. Finally, we shall examine the derivative of the phase---essentially the chirp. Recall in Chapter \ref{chap:linear} the chirp was defined as coefficient of $-\frac{1}{2}iT^2$ in the exponential, by taking the negative of the derivative we would expect a linear function with a slope equal to the chirp. In the nonlinear case, this is what we find for moderate values of $T$. On the other hand, for $|T| > 1$ the relation begins to level off. This is in agreement with experimental results \cite{chen, rothenberg, tomlinson}. \\

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\input{./Figures/StabilityZoom}
\caption{Centralized for typical $s$, and $b$ values.}
\label{fig:energyzoom}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\input{./Figures/Stability}
\caption{Energy of the pulse within the $s$-$b$ plane.}
\label{fig:energybig}
\end{subfigure}
\caption[Energy of the pulse at equilibrium.]{Energy of the pulse at equilibrium. The energy is constant along the black lines.}
\label{fig:energy}
\end{figure}

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_Shape}%
\input{./Figures/Unstable_Bad_Shape}
\caption{Envelope}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_FT}%
\input{./Figures/Unstable_Bad_FT}
\caption{Fourier transform}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\centering
\input{./Figures/Unstable_Chirp}%
\input{./Figures/Unstable_Bad_Chirp}
\caption{Chirp}
\end{subfigure}
\caption[Envelope, Fourier transform, and chirp of the pulse---unstable case.]{$s = 0.15$, $b = 2.15$, \textbf{Left:} 11 loops, \textbf{Right:} 16 loops}
\label{fig:break}
\end{figure}

@@

The noise in the energy exhibited for moderate to large values of $b$, and small values of $s$ is a phenomenon called \emph{wave breaking} \cite{agrawal2013, anderson, finot, rothenberg, tomlinson}. Wave breaking is not limited to just optics, wave breaking occurs in areas such as plasmas, transmission lines, and fluid dynamics \cite{rothenberg}. Wave breaking occurs because the pulse begins to interfere with itself in a way called self-phase modulation \cite{agrawal2002, agrawal2013, becker}. Self-phase modulation occurs because the index of refraction is intensity dependent \cite{agrawal2002, becker, rothenberg, silfvast}, which leads to additional chirp across the pulse \cite{agrawal2013, anderson, rothenberg, silfvast}. This in turn causes higher order frequencies to be injected into the pulse \cite{agrawal2013, anderson}. These high frequencies compound with each trip around the cavity becoming parasitic very quickly---Figure \ref{fig:break} highlights this. Notice that the difference between this figure and Figure \ref{fig:nlstable} is a difference in $b$ of $0.05$---this difference could be as small as adding a few centimetres more of fibre between the gain and output coupler. The left figures show the pulse after 11 trips around the cavity, in the Fourier transform it is clear that the contributions from higher frequencies has increased---we obtain similar results as in \cite{anderson, rothenberg}. Additionally, the chirp starts losing its linearity causing it to start becoming unstable; the nature of this instability is in agreement with \cite{anderson, rothenberg}. The parasitic nature of the high frequency contributions is evident by examining the right figures. After 5 additional trips around the cavity, the envelope of the pulse is much more rippled, and the real and imaginary parts become incoherent. Moreover, the Fourier transform has no clear structure and has essentially become noise, and the chirp has grown to be highly oscillatory and unstable. Once the pulse has reached a state such as this, there is no turning back---the envelope, Fourier transform, and chirp never reach a steady equilibrium state. \\

\subsection{Convergence}
\begin{figure}[tbp]
\input{./Figures/Step1}
\caption[Error of the pulse.]{Error of the pulse (given by \eqref{eq:error}) between iterations $99$ and $100$.}
\label{fig:error}
\end{figure}

\begin{figure}[p]
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step2}
\caption{$\Delta = 2$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step3}
\caption{$\Delta = 3$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step5}
\caption{$\Delta = 5$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step7}
\caption{$\Delta = 7$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step11}
\caption{$\Delta = 11$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step13}
\caption{$\Delta = 13$}
\end{subfigure} \\
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step16}
\caption{$\Delta = 16$}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
\input{./Figures/Step60}
\caption{$\Delta = 60$}
\end{subfigure} \\
\caption[Error for various values of $\Delta$.]{Error for various values of $\Delta$, the colour bar and axis labels have been omitted for clarity, the scales and ranges are the same as in Figure \ref{fig:error}.}
\label{fig:deltaerror}
\end{figure}

@@

The second cause of these errors is because some equilibrium states have a periodicity greater than $1$, we redefine \eqref{eq:error} to be instead
\begin{align}
\textrm{E}_\Delta = \frac{\| |A_i| - |A_{i-\Delta}| \|_2}{\| A_{i-\Delta} \|_2},
\label{eq:deltaerror}
\end{align}
to pick up equilibrium states with period $\Delta$, again with $i = 100$. Figure \ref{fig:deltaerror} shows the error for an assortment of $\Delta$ values. There are several key observations to make, first, for $\Delta = 2$ the kidney-shaped blob in the lower right corner has vanished, thus, this entire region has a periodicity of $2$. Second, for $\Delta = 3$ in the upper right corner a negative triangular-shaped structure emerges. Lastly, $\Delta \in \left\{ 5, 7, 11, 13 \right\}$ may not seem to extract any higher order periods, however, there are in fact very small striations in the lower left. So far, the $\Delta$ values have been prime since they are the building blocks for composite numbers. However, there are two composite $\Delta$ values of interest---the first being $16$. Since $16 = 2^4$ it will pick up any periodic behaviour with orders of $2, 4, 8, 16$, which are frequently found in bifurcations. Lastly, we chose $\Delta = 60$ for a similar reason: because $60$ is highly divisible---it is an abundant number. This extracts orders such as $12$ or $30$ without having to explicitly run the calculations\footnote{In the case of $\Delta = 60$ a value of $i = 150$ is used to ensure sufficient iterations to converge.}. \\

\begin{figure}[tbp]
\input{./Figures/Min}
\caption[Composite error.]{Composite error, \eqref{eq:errorcomp}, of Figure \ref{fig:deltaerror}.}
\label{fig:errorcomp}
\end{figure}

@@

\begin{figure}[p]
\centering
\begin{subfigure}{\textwidth}
\input{./Figures/StabilitySwitchZoom}
\caption{Centralized for typical $s$, and $b$ values.}
\label{fig:switchzoom}
\end{subfigure} \\
\begin{subfigure}{\textwidth}
\input{./Figures/StabilitySwitch}
\caption{Energy of the pulse within the $s$-$b$ plane.}
\label{fig:switchbig}
\end{subfigure}
\caption{Energy of the pulse at equilibrium, with the modulation and dispersion blocks switched.}
\label{fig:switch}
\end{figure}

### 2019-05-22 Thesis/Previous_Models/Previous.tex
Made some edits.

\chapter{Previous Modelling Efforts}
In this chapter we shall explore the previous modelling efforts by first reviewing the classic equations used in nonlinear optics. We will then build upon this to the master equation of mode-locking. The solutions to this will briefly be studied before diving into `discretized' functional models. \\

\section{Generalized Nonlinear Schr\"odinger Equation}
The standard equation for studying nonlinear optics is the nonlinear Schr\"odinger equation (NLSE) \cite{agrawal2013, anderson, burgoyne2007, desurvire, ferreira, finot, rothenberg}
\begin{align*}
\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
\end{align*}
Here, $A$ is the complex pulse amplitude, $\beta_2$ is the second-order---or group delay---dispersion, and $\gamma$ is the coefficient of nonlinearity. In practice, this equation lacks a few key terms. Thus, it is often generalized by adding amplification, loss, and higher order terms. This gives the generalized nonlinear Schr\"odinger equation (GNLSE) \cite{agrawal2013, bohun, finot, peng, shtyrina, yarutkina},
\begin{align}
\label{eq:nlse}
\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + \frac{\beta_3}{6}\pdiff[3]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
\end{align}
where $\beta_3$ is the third order effect of dispersion, $g(A)$ is an amplifying term due to the gain, and $\alpha$ is the loss due to scattering and absorption. \\

@@

In the least complicated case, the modulation and nonlinearity are omitted resulting in a solution in the form of a hyperbolic secant \cite{haus1975, haus1986, haus1992}. Now, with the inclusion of the nonlinearity a similar solution is found, however, it is now it the form of a chirped hyperbolic secant \cite{haus1991, usechak}. Both of these are unsurprising since this reduces to the soliton solution of the NLSE \cite{ferreira}. Lastly, by including the modulation term, and excluding the nonlinearity, \eqref{eq:meml} can be solved using separation of variables and one finds the solutions are the Gaussian--Hermite polynomials \cite{burgoyne2014, hausbook, haus1975, haus1996, haus2000, kartner, tamura, usechak}. However, in practice only the Gaussian is stable---the higher modes quickly decay within the laser \cite{hausbook, haus1975, haus1996, haus2000}. For a more comprehensive and exhaustive history see \cite{haus2000}. \\

\section{Discrete Functional Model}
While solution to the master equation yields reasonable results, it is not necessarily representative of what happens within the laser cavity. The issue with \eqref{eq:meml} is that it assumes each process effects the pulse continuously within the cavity. As highlighted by Figure \ref{fig:cavity}, this is a poor assumption. Within the cavity each effect is localized to its corresponding component: almost all of the dispersion happens within the CFBG, the pulse is only amplified within the Erbium-doped fibre, etc. Thus, perhaps a better model is one where \eqref{eq:meml} is broken down into the individual components giving the effect of each `block' of the cavity. Each of the blocks can then be composed together functionally to give the effect of one circuit around the cavity. This yields an algebraic equation instead of a differential one. \\

Such a method was first proposed in 1955 by Cutler \cite{cutler} while analyzing a microwave regenerative pulse generator. This method was adapted for mode-locked lasers in 1969 by Siegman and Kuizenga \cite{siegman}, which they then greatly added to the following year \cite{kuizenga1970a}. Kuizenga and Seigman also had success experimentally validating their model \cite{kuizenga1970b, kuizenga1970}. The effects of the nonlinearity would not be considered until Martinez, Fork, and Gordon \cite{martinez1984, martinez1985} tried to model passive mode-locking---mode-locking without the use of a modulator. In the absence of a modulator the nonlinearity becomes crucial to shaping the pulse. I don't know how to start this sentence*** when it was resurrected by Burgoyne \cite{burgoyne2014} in 2014 for tuneable lasers. \\

Despite these attempts, several short-comings exist. The clearest is that none of these models have contained each block--either the nonlinearity or the modulation have been omitted. In the context of tuneable lasers each component plays an important role. Another key drawback is that the functional operations of some of the components are somewhat phenomenological. While these functions are chosen based on the observed output, they are not necessarily correct. Finally, none of these previous models have been able to exhibit a phenomena called \emph{wave-breaking} in which the self-phase modulation (SPM) of the pulse becomes too strong distorting and damaging the wave until it ultimately becomes unstable and unsustainable. This notion will be explained in greater detail in Chapter \ref{chap:nl}. \\

>>>>

\chapter{Previous Modelling Efforts}
In this chapter we shall explore the previous modelling efforts by first reviewing the classic equations used in nonlinear optics. We will then build upon this to the master equation of mode-locking, where the solutions to this will briefly be studied before diving into `discretized' functional models. \\

\section{Generalized Nonlinear Schr\"odinger Equation}
The standard equation for studying nonlinear optics is the nonlinear Schr\"odinger equation (NLSE) \cite{agrawal2013, anderson, burgoyne2007, desurvire, ferreira, finot, rothenberg}
\begin{align*}
\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
\end{align*}
Here $A$ is the complex pulse amplitude, $\beta_2$ is the second-order---or group delay---dispersion, and $\gamma$ is the coefficient of nonlinearity or Kerr coefficient. In practice, this equation lacks a few key terms. Thus, it is often generalized by adding amplification, loss, and higher order terms. This gives the generalized nonlinear Schr\"odinger equation (GNLSE) \cite{agrawal2013, bohun, finot, peng, shtyrina, yarutkina},
\begin{align}
\label{eq:nlse}
\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + \frac{\beta_3}{6}\pdiff[3]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
\end{align}
where $\beta_3$ is the third order coefficient of dispersion, $g(A)$ is an amplifying term due to the gain, and $\alpha$ is the loss due to scattering and absorption. \\

@@

In the least complicated case, the modulation and nonlinearity are omitted resulting in a solution in the form of a hyperbolic secant \cite{haus1975, haus1986, haus1992}. With the inclusion of the nonlinearity a similar solution is found, however, it is now it the form of a chirped hyperbolic secant \cite{haus1991, usechak}. Both of these are unsurprising since this reduces to the soliton solution of the NLSE \cite{ferreira}. Lastly, by including the modulation term, and excluding the nonlinearity, \eqref{eq:meml} can be solved using separation of variables and one finds the solutions are the Gaussian--Hermite polynomials \cite{burgoyne2014, hausbook, haus1975, haus1996, haus2000, kartner, tamura, usechak}. However, in practice only the Gaussian is stable---the higher modes quickly decay within the laser \cite{hausbook, haus1975, haus1996, haus2000}. For a more comprehensive and exhaustive history see \cite{haus2000}. \\

\section{Discrete Functional Model}
While solutions to the master equation yield reasonable results, it is not necessarily representative of what happens within the laser cavity. The issue with \eqref{eq:meml} is that it assumes each process effects the pulse continuously within the cavity. As highlighted by Figure \ref{fig:cavity}, this is a rather poor assumption. Within the cavity each effect is localized to its corresponding component: almost all of the dispersion happens within the CFBG, the pulse is only amplified within the Erbium-doped fibre, etc. Thus, perhaps a better model is one where \eqref{eq:meml} is broken down into the individual components giving the effect of each `block' of the cavity. Each of the blocks can then be composed together functionally to give the effect of one circuit around the cavity. This yields an algebraic equation instead of a differential one. \\

Such a method was first proposed in 1955 by Cutler \cite{cutler} while analyzing a microwave regenerative pulse generator. This method was adapted for mode-locked lasers in 1969 by Siegman and Kuizenga \cite{seigman}, which they then greatly added to the following year \cite{kuizenga1970a}. Kuizenga and Seigman also had success experimentally validating their model \cite{kuizenga1970b, kuizenga1970}. The effects of the nonlinearity would not be considered until Martinez, Fork, and Gordon \cite{martinez1984, martinez1985} tried to model passive mode-locking---mode-locking without the use of a modulator. In the absence of a modulator the nonlinearity becomes crucial to shaping the pulse. I don't know how to start this sentence*** when it was resurrected by Burgoyne \cite{burgoyne2014} in 2014 for tuneable lasers. \\

Despite these attempts, several short-comings exist. The clearest is that none of these models have contained every block--either the nonlinearity or the modulation have been omitted. In the framework of tuneable lasers each component plays an important role. Another key drawback is that the functional operations of some of the components are somewhat phenomenological. While these functions are chosen based on the observed output, they are not necessarily correct. Finally, none of these previous models have been able to exhibit a phenomena called \emph{wave-breaking} in which the self-phase modulation (SPM) of the pulse becomes too strong distorting and damaging the wave until it ultimately becomes unstable and unsustainable. This notion will be explained in greater detail in Chapter \ref{chap:nl}. \\
